 ---
 title: "Main functionality"
 author: "Lukas Tilmann"
 date: "`r Sys.Date()`"
 output: rmarkdown::html_vignette
 vignette: >
   %\VignetteIndexEntry{Main functionality}
   %\VignetteEngine{knitr::rmarkdown}
   %\VignetteEncoding{UTF-8}
   %\VignetteDepends{tf, ggplot2}
 ---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(PeakPersistance)
library(tf)
library(tidyfun)
library(ggplot2)
```

```{r}
combined_func <- generate_benchmark_function(
  bases = list(3, "normal", "normal"), 
  coefficients = list(
    2.5,
    10,
    10
  ),
  shifts = c(0, 0.3, 0.7),     # Shifts for each basis
  scale_factors = c(1, 10, 10)      # Amplitude factors
)

curve_data <- generate_functional_curves(n = 25,
                                         num_points = 100,
                                         g = combined_func,
                                         warping = "simple",
                                         sigma_amplitude = 0.01,
                                         noise_to_signal = 0.002,
                                         seed = 123)


plot_simulated_curves(curves = curve_data$curves,
                      t_grid = curve_data$t_grid,
                      g = curve_data$base_function)
```
```{r}
tau <- find_tau(curve_data$curves, curve_data$t_grid, 15)
tau
```


```{r}
t_grid <- curve_data$t_grid
curves <- curve_data$curves

n_lambda <- 5
lam_start <- 2.5
sig_threshold <- tau
pers_threshold <- 0.28
```


```{r}
ret <- peak_persistance_diagram(curves = curves, 
                                t_grid = t_grid,
                                parallel = FALSE,
                                lambda_search_start = lam_start,
                                lambda_search_min_bound = 0.1,
                                lambda_search_threshold = 1e-3,
                                max_lambda_search_steps = 5,
                                n_lambda = n_lambda,
                                max_iter = 5,
                                sig_threshold = sig_threshold, 
                                pers_threshold = pers_threshold,
                                pers_method = "clustering")
```

Peak-persistance diagram in barchart form
```{r}
ret$bc
```

Peak-persistance diagram in heatmap form
```{r}
ret$surface
```

Shape constrained density estimation

```{r}
mfn <- ret$mean_function
peak_locs <- ret$peak_locs
valley_locs <- ret$valley_locs
significant_peaks <- ret$significant_peaks
peak_labels <- ret$labels
t_grid <- ret$time_grid
fun_aligned <- ret$aligned_functions

k = 8
rho = 1e-6
basis_type = "tp"

fn_est <- shape_constrained_estimation(fun_aligned, peak_locs, valley_locs, 
                                       significant_peaks, peak_labels, mfn, k, 
                                       t_grid, rho, basis_type)
```



```{r}
vis_df <- data.frame(curves_aligned = fun_aligned, curves_unaligned = curves)
ground_truth <- tfd(curve_data$base_function(t_grid), t_grid)

ggplot(vis_df, aes(y = fun_aligned)) +
  geom_spaghetti(alpha = 0.3) +
  geom_spaghetti(aes(y = fn_est), color = "orange", linewidth = 1.5) +
  geom_spaghetti(aes(y = mfn), color = "green", linewidth = 1.5) +
  geom_spaghetti(aes(y = ground_truth), color = "red", linewidth = 1.5)
```


More complex simulated data
```{r}
combined_func_complex <- generate_benchmark_function(
  bases = list(3, "normal", "normal"),  # x^3, normal, normal, sine
  coefficients = list(
    1,
    10,
    3
  ),
  shifts = c(0.5, 0.3, 0.7),     # Shifts for each basis
  scale_factors = c(2,10, 10)      # Amplitude factors
)

curve_data_complex <- generate_functional_curves(n = 50,
                                         num_points = 100,
                                         g = combined_func_complex,
                                         warping = "flexible",
                                         warping_points = 5,
                                         warping_gamma = 5,
                                         sigma_amplitude = 0.1,
                                         noise_to_signal = 0.002,
                                         seed = 123)


plot_simulated_curves(curves = curve_data_complex$curves,
                      t_grid = curve_data_complex$t_grid,
                      g = curve_data_complex$base_function)
```
```{r}
tau <- find_tau(curve_data_complex$curves, curve_data_complex$t_grid, 25)
tau
```


```{r}
t_grid_c <- curve_data$t_grid
curves_c <- curve_data_complex$curves

n_lambda <- 15
lam_start <- 2
sig_threshold <- tau
pers_threshold <- 0.28
ret_c <- peak_persistance_diagram(curves = curves_c, 
                                t_grid = t_grid,
                                parallel = FALSE,
                                lambda_search_start = lam_start,
                                lambda_search_min_bound = 0.01,
                                lambda_search_threshold = 1e-3,
                                n_lambda = 25,
                                sig_threshold = sig_threshold, 
                                pers_threshold = pers_threshold)
```


```{r}
ret_c$surface
```

```{r}
mfn_c <- ret_c$mean_function
peak_locs_c <- ret_c$peak_locs
valley_locs_c <- ret_c$valley_locs
significant_peaks_c <- ret_c$significant_peaks
peak_labels_c <- ret_c$labels
t_grid_c <- ret_c$time_grid
fun_aligned_c <- ret_c$aligned_functions

k = 8
rho = 1e-7
basis_type = "tp"

fn_est_c <- shape_constrained_estimation(fun_aligned_c, peak_locs_c, valley_locs_c, 
                                       significant_peaks_c, peak_labels_c, mfn_c, k, 
                                       t_grid_c, rho, basis_type)
```


```{r}
ground_truth_c <- tf::tfd(curve_data_complex$base_function(t_grid_c), t_grid_c)
vis_df_c <- data.frame(curves_aligned = fun_aligned_c)
ggplot(vis_df_c, aes(y = fun_aligned_c)) +
  geom_spaghetti(alpha = 0.3) +
  geom_spaghetti(aes(y = fn_est_c), color = "green", linewidth = 1.5) + 
  geom_spaghetti(aes(y = mfn_c), color = "orange", linewidth = 1.5) +
  geom_spaghetti(aes(y = ground_truth_c), color = "red", linewidth = 1.5)
  
```

```{r}
step_2_align <- align_to_mean(curves_c, fn_est_c, t_grid_c, 0)
plot(step_2_align$aligned_functions, t_grid)
```

