---
title: "study_1_analysis_prelim"
output: html_document
---

# Benchmark Study Preliminary Analysis

This analysis explores impact of various data and model parameters on quality of results.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(knitr)
library(broom)
library(patchwork)
library(tidyfun)
library(kableExtra)
library(gridExtra)
```

## Data Loading and Preparation

```{r load-data}
# Load benchmark data (adjust path as needed)
benchmark_data <- read.csv("../data/metrics_dataframe_study_1.csv")


# Convert all model and data parameters to factors
benchmark_data$g_id <- as.factor(benchmark_data$g_id)
benchmark_data$warping <- as.factor(benchmark_data$warping)
benchmark_data$penalty <- as.factor(benchmark_data$penalty)
benchmark_data$lambda_grid_spacing <- as.factor(benchmark_data$lambda_grid_spacing)
benchmark_data$n <- as.factor(benchmark_data$n)
benchmark_data$noise_to_signal <- as.factor(benchmark_data$noise_to_signal)
benchmark_data$sigma_amplitude <- as.factor(benchmark_data$sigma_amplitude)
benchmark_data$warping_gamma <- as.factor(benchmark_data$warping_gamma)
benchmark_data$warping_points <- as.factor(benchmark_data$warping_points)
benchmark_data$curvature_percentile <- as.factor(benchmark_data$curvature_percentile)
```

Distances need to be normalized as functions have different variances. Also computing correct number of peaks for functions.
```{r pre-processing}
devtools::load_all(".")

# Define the template functions as they were in the benchmark study
fun_1 <- generate_benchmark_function(list("normal", "normal"),
                                     c(10, 10), c(0.3, 0.7), c(10, 10))

fun_2 <- generate_benchmark_function(
  bases = list("normal", "normal", "normal", "normal"),  
  coefficients = list(
    10,
    5,
    3,
    7# Coefficient for sine
  ),
  shifts = c(0.2, 0.5, 0.7, 0.9),     # Shifts for each basis
  scale_factors = c(15, 10, 20, 10)      # Amplitude factors
)

fun_3 <- generate_benchmark_function(
  bases = list("sine", "normal", "normal", "normal", "normal"),
  coefficients = list(
    10,
    -30,
    -30,
    -30,
    -30
  ),
  shifts = c(0.8, 0.09, 0.35, 0.6, 0.85),
  scale_factors = c(25, 75, 75, 75, 75)
)

# Create a grid for evaluation
t_grid <- seq(0, 1, length.out = 100)

# Convert to tidyfun objects, similar to how they're used in the benchmark
f1_tf <- tfd(fun_1(t_grid), arg = t_grid)
f2_tf <- tfd(fun_2(t_grid), arg = t_grid)
f3_tf <- tfd(fun_3(t_grid), arg = t_grid)

# Calculate variance using tf_fvar (functional variance) as used in the benchmark
f1_var <- tf_fvar(f1_tf, t_grid)
f2_var <- tf_fvar(f2_tf, t_grid)
f3_var <- tf_fvar(f3_tf, t_grid)

# Store variances in a named vector for easy lookup
function_variances <- c(
  "1" = f1_var,
  "2" = f2_var,
  "3" = f3_var
)

# Find peaks using the same methodology as analyze_mean_function in the benchmark
count_peaks <- function(f_tf) {
  # Calculate derivatives
  slope <- tf_derive(f_tf, order = 1)
  
  # Find peaks (local maxima) - where slope changes from positive to negative
  peaks <- tf_where(
    slope,
    value < 0 & dplyr::lag(value, 1, value[1]) > 0)
  
  # Return number of peaks and their locations
  return(length(peaks[[1]]))
}

# Count peaks for each function
f1_num_peaks <- count_peaks(f1_tf)
f2_num_peaks <- count_peaks(f2_tf)
f3_num_peaks <- count_peaks(f3_tf)

# Store peak counts in a named vector for easy lookup
function_peaks <- c(
  "1" = f1_num_peaks,
  "2" = f2_num_peaks,
  "3" = f3_num_peaks
)
```

```{r pre-processing-2}
# Add normalized columns to the benchmark data
benchmark_data <- benchmark_data %>%
  mutate(
    # Convert g_id to character for lookup
    g_id_chr = as.character(g_id),
    
    # Create normalized distance metrics
    normalized_mean_distance = mean_ground_truth_distance / sqrt(function_variances[g_id_chr]),
    normalized_est_distance = est_ground_truth_distance / sqrt(function_variances[g_id_chr]),
    
    # Calculate peak detection error
    true_num_peaks = function_peaks[g_id_chr],
    peak_detection_error = abs(num_peaks - true_num_peaks)
  )

# Convert g_id to factor for better analysis
benchmark_data$g_id <- as.factor(benchmark_data$g_id)

# Preview the new columns
head(benchmark_data %>% select(g_id, mean_ground_truth_distance, normalized_mean_distance, 
                              est_ground_truth_distance, normalized_est_distance,
                              num_peaks, true_num_peaks, peak_detection_error))


```

```{r}
# Preview the data
head(benchmark_data, n = 25) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                font_size = 11)

```


## 1. Visualizing the Impact of Data Parameters

### 1.1 Function Type (g_id)

```{r g-id-impact}
# Plot performance metrics by function type
p1 <- ggplot(benchmark_data, aes(x = g_id, y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = g_id), alpha = 0.7) +
  labs(title = "Normalized Mean Function Distance by Function Type",
       x = "Function Type (g_id)", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = g_id, y = normalized_est_distance)) +
  geom_boxplot(aes(fill = g_id), alpha = 0.7) +
  labs(title = "Normalized Estimated Function Distance by Function Type",
       x = "Function Type (g_id)", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = g_id, y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = g_id), alpha = 0.7) +
  labs(title = "Warping Distance by Function Type",
       x = "Function Type (g_id)", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = g_id, y = peak_detection_error)) +
  geom_boxplot(aes(fill = g_id), alpha = 0.7) +
  labs(title = "Peak Detection Error by Function Type",
       x = "Function Type (g_id)", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots in a 2x2 grid
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### 1.2 Sample Size (n)

```{r n-impact}
# Plot performance metrics by sample size
p1 <- ggplot(benchmark_data, aes(x = n, y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(n)), alpha = 0.7) +
  labs(title = "Normalized Mean Distance by Sample Size",
       x = "Sample Size (n)", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = n, y = normalized_est_distance)) +
  geom_boxplot(aes(fill = as.factor(n)), alpha = 0.7) +
  labs(title = "Normalized Estimated Distance by Sample Size",
       x = "Sample Size (n)", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = n, y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(n)), alpha = 0.7) +
  labs(title = "Warping Distance by Sample Size",
       x = "Sample Size (n)", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = n, y = peak_detection_error)) +
  geom_boxplot(aes(fill = as.factor(n)), alpha = 0.7) +
  labs(title = "Peak Detection Error by Sample Size",
       x = "Sample Size (n)", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### 1.3 Noise Level (noise_to_signal)

```{r noise-impact}
# Plot performance metrics by noise level
p1 <- ggplot(benchmark_data, aes(x = as.factor(noise_to_signal), y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(noise_to_signal)), alpha = 0.7) +
  labs(title = "Normalized Mean Distance by Noise Level",
       x = "Noise to Signal Ratio", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = as.factor(noise_to_signal), y = normalized_est_distance)) +
  geom_boxplot(aes(fill = as.factor(noise_to_signal)), alpha = 0.7) +
  labs(title = "Normalized Estimated Distance by Noise Level",
       x = "Noise to Signal Ratio", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = as.factor(noise_to_signal), y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(noise_to_signal)), alpha = 0.7) +
  labs(title = "Warping Distance by Noise Level",
       x = "Noise to Signal Ratio", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = as.factor(noise_to_signal), y = peak_detection_error)) +
  geom_boxplot(aes(fill = as.factor(noise_to_signal)), alpha = 0.7) +
  labs(title = "Peak Detection Error by Noise Level",
       x = "Noise to Signal Ratio", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### 1.4 Amplitude Variation (sigma_amplitude)

```{r amplitude-impact}
# Plot performance metrics by amplitude variation
p1 <- ggplot(benchmark_data, aes(x = as.factor(sigma_amplitude), y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(sigma_amplitude)), alpha = 0.7) +
  labs(title = "Normalized Mean Distance by Amplitude Variation",
       x = "Amplitude Variation (sigma)", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = as.factor(sigma_amplitude), y = normalized_est_distance)) +
  geom_boxplot(aes(fill = as.factor(sigma_amplitude)), alpha = 0.7) +
  labs(title = "Normalized Estimated Distance by Amplitude Variation",
       x = "Amplitude Variation (sigma)", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = as.factor(sigma_amplitude), y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(sigma_amplitude)), alpha = 0.7) +
  labs(title = "Warping Distance by Amplitude Variation",
       x = "Amplitude Variation (sigma)", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = as.factor(sigma_amplitude), y = peak_detection_error)) +
  geom_boxplot(aes(fill = as.factor(sigma_amplitude)), alpha = 0.7) +
  labs(title = "Peak Detection Error by Amplitude Variation",
       x = "Amplitude Variation (sigma)", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

### 1.5 Warping Method

```{r warping-impact}
# Plot performance metrics by warping method
p1 <- ggplot(benchmark_data, aes(x = warping, y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = warping), alpha = 0.7) +
  labs(title = "Normalized Mean Distance by Warping Method",
       x = "Warping Method", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = warping, y = normalized_est_distance)) +
  geom_boxplot(aes(fill = warping), alpha = 0.7) +
  labs(title = "Normalized Estimated Distance by Warping Method",
       x = "Warping Method", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = warping, y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = warping), alpha = 0.7) +
  labs(title = "Warping Distance by Warping Method",
       x = "Warping Method", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = warping, y = peak_detection_error)) +
  geom_boxplot(aes(fill = warping), alpha = 0.7) +
  labs(title = "Peak Detection Error by Warping Method",
       x = "Warping Method", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

## 2. Impact of Model Parameters - Same Analysis for Model Parameters

### 2.1 Penalty

```{r penalty-impact}
# Plot performance metrics by penalty type
p1 <- ggplot(benchmark_data, aes(x = penalty, y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = penalty), alpha = 0.7) +
  labs(title = "Normalized Mean Distance by Penalty Type",
       x = "Penalty Type", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = penalty, y = normalized_est_distance)) +
  geom_boxplot(aes(fill = penalty), alpha = 0.7) +
  labs(title = "Normalized Estimated Distance by Penalty Type",
       x = "Penalty Type", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = penalty, y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = penalty), alpha = 0.7) +
  labs(title = "Warping Distance by Penalty Type",
       x = "Penalty Type", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = penalty, y = peak_detection_error)) +
  geom_boxplot(aes(fill = penalty), alpha = 0.7) +
  labs(title = "Peak Detection Error by Penalty Type",
       x = "Penalty Type", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

```

### 2.2 Lambda Grid Spacing
```{r spacing-impact}

# Plot performance metrics by lambda grid spacing
p1 <- ggplot(benchmark_data, aes(x = lambda_grid_spacing, y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = lambda_grid_spacing), alpha = 0.7) +
  labs(title = "Normalized Mean Distance by Lambda Grid Spacing",
       x = "Lambda Grid Spacing", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = lambda_grid_spacing, y = normalized_est_distance)) +
  geom_boxplot(aes(fill = lambda_grid_spacing), alpha = 0.7) +
  labs(title = "Normalized Estimated Distance by Lambda Grid Spacing",
       x = "Lambda Grid Spacing", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = lambda_grid_spacing, y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = lambda_grid_spacing), alpha = 0.7) +
  labs(title = "Warping Distance by Lambda Grid Spacing",
       x = "Lambda Grid Spacing", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = lambda_grid_spacing, y = peak_detection_error)) +
  geom_boxplot(aes(fill = lambda_grid_spacing), alpha = 0.7) +
  labs(title = "Peak Detection Error by Lambda Grid Spacing",
       x = "Lambda Grid Spacing", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

```

### 2.3 Curvature Percentile
```{r curvature-impact}
# Plot performance metrics by curvature percentile
p1 <- ggplot(benchmark_data, aes(x = as.factor(curvature_percentile), y = normalized_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(curvature_percentile)), alpha = 0.7) +
  labs(title = "Normalized Mean Distance by Curvature Percentile",
       x = "Curvature Percentile", 
       y = "Normalized Mean Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(benchmark_data, aes(x = as.factor(curvature_percentile), y = normalized_est_distance)) +
  geom_boxplot(aes(fill = as.factor(curvature_percentile)), alpha = 0.7) +
  labs(title = "Normalized Estimated Distance by Curvature Percentile",
       x = "Curvature Percentile", 
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- ggplot(benchmark_data, aes(x = as.factor(curvature_percentile), y = warping_distances_ppd_warping_mean_distance)) +
  geom_boxplot(aes(fill = as.factor(curvature_percentile)), alpha = 0.7) +
  labs(title = "Warping Distance by Curvature Percentile",
       x = "Curvature Percentile", 
       y = "Warping Distance") +
  theme_minimal() +
  theme(legend.position = "none")

p4 <- ggplot(benchmark_data, aes(x = as.factor(curvature_percentile), y = peak_detection_error)) +
  geom_boxplot(aes(fill = as.factor(curvature_percentile)), alpha = 0.7) +
  labs(title = "Peak Detection Error by Curvature Percentile",
       x = "Curvature Percentile", 
       y = "Peak Detection Error") +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

```



## 3. Regression Models for Impact of Variables on Various Metrics

### Model 1: Mean Function Distance from Original Function
```{r model-mean-distance}
model_norm_mean <- lm(normalized_mean_distance ~ g_id + n + noise_to_signal + 
                      sigma_amplitude + warping + penalty + lambda_grid_spacing, 
                      data = benchmark_data)

# Model summary
summary_norm_mean <- summary(model_norm_mean)

# Tidy model output for better display
tidy_norm_mean <- tidy(model_norm_mean, conf.int = TRUE)

# Display results
kable(tidy_norm_mean, caption = "Regression Model for Normalized Mean Distance", digits = 4)
```

### Model 2: Estimated function Distance from Original Function

```{r model-est-distance}
# Fit model for normalized estimated distance
model_norm_est <- lm(normalized_est_distance ~ g_id + n + noise_to_signal + 
                     sigma_amplitude + warping + penalty + lambda_grid_spacing, 
                     data = benchmark_data)

# Model summary
summary_norm_est <- summary(model_norm_est)

# Tidy model output for better display
tidy_norm_est <- tidy(model_norm_est, conf.int = TRUE)

# Display results
kable(tidy_norm_est, caption = "Regression Model for Normalized Estimated Distance", digits = 4)
```

### Model 3: Distance Between Warping Functions from Peak-Persistence-Diagram and Original Warping Functions

```{r model-warping-distance}
# Fit model for warping mean distance
model_warp <- lm(warping_distances_ppd_warping_mean_distance ~ g_id + n + noise_to_signal + 
                 sigma_amplitude + warping + penalty + lambda_grid_spacing, 
                 data = benchmark_data)

# Model summary
summary_warp <- summary(model_warp)

# Tidy model output for better display
tidy_warp <- tidy(model_warp, conf.int = TRUE)

# Display results
kable(tidy_warp, caption = "Regression Model for Warping Mean Distance", digits = 4)
```


### Model 4: Model with Interaction Effects between Model and Data Parameters for 
###          Distance between Mean and Original Function

```{r interactions-model}

# Model with interactions between data and model params for normalized mean distance
model_norm_mean_int <- lm(normalized_mean_distance ~ 
                          # Main effects for all parameters
                          g_id + n + noise_to_signal + sigma_amplitude + 
                          warping + penalty + lambda_grid_spacing + curvature_percentile +
                          # Interactions between data and model parameters
                          g_id:(penalty + lambda_grid_spacing + curvature_percentile) +
                          n:(penalty + lambda_grid_spacing + curvature_percentile) +
                          noise_to_signal:(penalty + lambda_grid_spacing + curvature_percentile) +
                          sigma_amplitude:(penalty + lambda_grid_spacing + curvature_percentile) +
                          warping:(penalty + lambda_grid_spacing + curvature_percentile),
                          data = benchmark_data)

# Compare models using ANOVA
anova_mean <- anova(model_norm_mean, model_norm_mean_int)

# Display ANOVA results
kable(anova_mean, caption = "ANOVA: Comparison of Mean Distance Models", digits = 4)

# Get top significant interactions for each model
top_interactions <- function(model, n = 10) {
  tidied <- tidy(model) %>%
    filter(grepl(":", term)) %>%
    arrange(p.value) %>%
    head(n)
  return(tidied)
}

top_int_mean <- top_interactions(model_norm_mean_int)

# Display top interactions
kable(top_int_mean, caption = "Top Significant Interactions - Mean Distance Model", digits = 4)
```

## 4. Analysis of Difference between Mean and Shape-constrained Estimate Function

```{r estimation-improvement}

# Compare normalized mean distance and normalized estimated distance
comparison_data <- benchmark_data %>%
  mutate(
    # Calculate change (positive values = worse performance)
    percent_change = (normalized_est_distance - normalized_mean_distance) / normalized_mean_distance
  )

# Summary statistics of change
change_summary <- comparison_data %>%
  summarize(
    mean_change = mean(percent_change, na.rm = TRUE),
    median_change = median(percent_change, na.rm = TRUE),
    min_change = min(percent_change, na.rm = TRUE),
    max_change = max(percent_change, na.rm = TRUE),
    sd_change = sd(percent_change, na.rm = TRUE)
  )

# Print summary table
kable(change_summary, caption = "Summary of Change from Mean Function Distance to Estimated Distance", 
      digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                font_size = 11)

# Plot distribution of percentage change
p_hist <- ggplot(comparison_data, aes(x = percent_change)) +
  geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Distribution of Change",
       x = "Change",
       y = "Count") +
  theme_minimal()

# Plot paired comparison
p_paired <- ggplot(comparison_data, aes(x = normalized_mean_distance, y = normalized_est_distance)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Paired Comparison of Normalized Distances",
       subtitle = "Points below the line indicate better performance with estimated function",
       x = "Normalized Mean Distance",
       y = "Normalized Estimated Distance") +
  theme_minimal() +
  coord_equal()

# Arrange the plots side by side
grid.arrange(p_hist, p_paired, ncol = 2)

# Perform paired t-test
t_test_result <- t.test(
  comparison_data$normalized_est_distance,
  comparison_data$normalized_mean_distance,
  paired = TRUE,
  alternative = "less"  # Test if estimated distance is significantly less than mean distance
)

# Print test results
cat("Paired t-test results:\n")
print(t_test_result)

# Perform non-parametric Wilcoxon signed-rank test (doesn't assume normality)
wilcox_test_result <- wilcox.test(
  comparison_data$normalized_est_distance,
  comparison_data$normalized_mean_distance,
  paired = TRUE,
  alternative = "less"
)

cat("\nWilcoxon signed-rank test results:\n")
print(wilcox_test_result)

# Analyze change by function type
change_by_function <- comparison_data %>%
  group_by(g_id) %>%
  summarize(
    mean_change = mean(percent_change, na.rm = TRUE),
    median_change = median(percent_change, na.rm = TRUE),
    n_samples = n()
  ) %>%
  arrange(mean_change)

kable(change_by_function, caption = "Change by Function Type (g_id)", 
      digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                font_size = 11)

# Box plot of change by function type
p_change_by_function <- ggplot(comparison_data, aes(x = g_id, y = percent_change, fill = g_id)) +
  geom_boxplot(alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(title = "Change by Function Type",
       x = "Function Type (g_id)",
       y = "Change") +
  theme_minimal() +
  theme(legend.position = "none")

print(p_change_by_function)

```

## 5. Drilldown on Effect of Model Parameters by Most Important Data Parameters (Function and Warping Type)


```{r}
# Create interaction plots for g_id and penalty
interaction_penalty_g_id <- ggplot(benchmark_data, aes(x = g_id, y = normalized_mean_distance, fill = penalty)) +
  geom_boxplot(position = position_dodge(width = 0.85), alpha = 0.7) +
  labs(
    title = "Interaction between Function Type and Penalty",
    x = "Function Type (g_id)",
    y = "Normalized Mean Distance",
    fill = "Penalty"
  ) +
  theme_minimal()

# Create interaction plots for warping and penalty
interaction_penalty_warping <- ggplot(benchmark_data, aes(x = warping, y = normalized_mean_distance, fill = penalty)) +
  geom_boxplot(position = position_dodge(width = 0.85), alpha = 0.7) +
  labs(
    title = "Interaction between Warping Type and Penalty",
    x = "Warping Type",
    y = "Normalized Mean Distance",
    fill = "Penalty"
  ) +
  theme_minimal()

# Create interaction plots for g_id and lambda_grid_spacing
interaction_spacing_g_id <- ggplot(benchmark_data, aes(x = g_id, y = normalized_mean_distance, fill = lambda_grid_spacing)) +
  geom_boxplot(position = position_dodge(width = 0.85), alpha = 0.7) +
  labs(
    title = "Interaction between Function Type and Lambda Grid Spacing",
    x = "Function Type (g_id)",
    y = "Normalized Mean Distance",
    fill = "Lambda Grid Spacing"
  ) +
  theme_minimal()

# Create interaction plots for warping and lambda_grid_spacing
interaction_spacing_warping <- ggplot(benchmark_data, aes(x = warping, y = normalized_mean_distance, fill = lambda_grid_spacing)) +
  geom_boxplot(position = position_dodge(width = 0.85), alpha = 0.7) +
  labs(
    title = "Interaction between Warping Type and Lambda Grid Spacing",
    x = "Warping Type",
    y = "Normalized Mean Distance",
    fill = "Lambda Grid Spacing"
  ) +
  theme_minimal()

# Create interaction plots for g_id and curvature_percentile
interaction_curvature_g_id <- ggplot(benchmark_data, aes(x = g_id, y = normalized_mean_distance, fill = curvature_percentile)) +
  geom_boxplot(position = position_dodge(width = 0.85), alpha = 0.7) +
  labs(
    title = "Interaction between Function Type and Curvature Percentile",
    x = "Function Type (g_id)",
    y = "Normalized Mean Distance",
    fill = "Curvature Percentile"
  ) +
  theme_minimal()

# Create interaction plots for warping and curvature_percentile
interaction_curvature_warping <- ggplot(benchmark_data, aes(x = warping, y = normalized_mean_distance, fill = curvature_percentile)) +
  geom_boxplot(position = position_dodge(width = 0.85), alpha = 0.7) +
  labs(
    title = "Interaction between Warping Type and Curvature Percentile",
    x = "Warping Type",
    y = "Normalized Mean Distance",
    fill = "Curvature Percentile"
  ) +
  theme_minimal()

# Print interaction plots
print(interaction_penalty_g_id)
print(interaction_penalty_warping)
print(interaction_spacing_g_id)
print(interaction_spacing_warping)
print(interaction_curvature_g_id)
print(interaction_curvature_warping)
```

